\documentclass[english,twoside,censored,tkt,sw-line]{HYthesisML}
\PassOptionsToClass{openright,twoside,a4paper}{report}
%\PassOptionsToClass{openany,twoside,a4paper}{report}
\usepackage{csquotes}
\usepackage{tikz}
\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\raggedbottom
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\usepackage{flafter}
\newcommand{\xmark}{\ding{55}}%
\usepackage[style=authoryear,bibstyle=authoryear,backend=biber,natbib=true,maxnames=99,maxcitenames=2,giveninits=true,uniquename=init]{biblatex}
\addbibresource{bibliography.bib}
\DeclareNameAlias{sortname}{family-given}
\usepackage{lmodern}         % Font package, again in some systems.
\usepackage{textcomp}        % Package for special symbols
\usepackage[euler]{textgreek}
\usepackage[pdftex]{color, graphicx} % For pdf output and jpg/png graphics
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage[pdftex, plainpages=false]{hyperref} % For hyperlinks and pdf metadata
\usepackage{fancyhdr}        % For nicer page headers
\usepackage{tikz}            % For making vector graphics (hard to learn but powerful)
%\usepackage{wrapfig}        % For nice text-wrapping figures (use at own discretion)
\usepackage{amsmath, amssymb} % For better math
\singlespacing               %line spacing options; normally use single
\fussy
\title{
Personalized Recommender Systems
        Bachelor Thesis}
\author{Anniina Sainio}
\date{May 2021}
\supervisors{Xiaoli Liu}
\examiners{Petteri Numri}
\keywords{Recommender Systems, Personalization, Unexpectedness, Feedback loop, Cold-start}

\additionalinformation{\translate{\track}}
\classification{\protect{\ \\
\  Computer systems organization $\rightarrow$ Architectures $\rightarrow$Other architectures$\rightarrow$ Neural networks\  \\
\  Information systems  $\rightarrow$  Information retrieval   $\rightarrow$ Retrieval tasks and goals  $\rightarrow$ Recommender systems
}}
\hypersetup{
    unicode=true,           % to show non-Latin characters in Acrobat’s bookmarks
    pdftoolbar=true,        % show Acrobat’s toolbar?
    pdfmenubar=true,        % show Acrobat’s menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={},            % title
    pdfauthor={},           % author
    pdfsubject={},          % subject of the document
    pdfcreator={},          % creator of the document
    pdfproducer={pdfLaTeX}, % producer of the document
    pdfkeywords={something} {something else}, % list of keywords for
    pdfnewwindow=true,      % links in new window
    colorlinks=true,        % false: boxed links; true: colored links
    linkcolor=black,        % color of internal links
    citecolor=black,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}

\begin{document}
\maketitle
\begin{otherlanguage}{english}
\begin{abstract}
Personalized Recommender Systems are designed to help users in decision making helping them choose items that meet their personal needs. This paper focuses on solutions to four common problems faced by current recommender systems and how five recent but different personalized recommender systems have solved or mitigated these problems. These four problems are: 1) The filter bubble problem causing boredom and dissatisfaction, 2) The cold-start problem, which is a condition where the circumstances  with new users and/or new items are unfavorable for the recommender system to yield the best possible results, 3) the problem of data sparsity when user-item historical data is sparse, and 4) the problem of dynamic, ever-changing user preferences. The Personalized Unexpected Recommender System (PURS) model, addresses the filter bubble problem effectively by integrating unexpectedness into the recommendation process. The Bayesian Graph Collaborative Filtering framework (BGCF), tackles the data sparsity issue utilizing a Bayesian Graph Neural Networks while also bringing about more diversity to the recommendations alleviating user boredom. The Graph-Regularized Cross-modal learning framework (GRC) undertakes the challenge of modeling complex non-linear user-item relations, exploiting both positive and negative user-item interactions. Moreover, most existing recommender systems use mostly historical data leading to cold-start problems for new users with no user history which the BGCF and GRC handle particularly well. The Deep Bayesian Bandits model focuses on resolving the feedback loop problem. The SSE-PT (Sequential Recommendation Via Personalized Transformer), is a neural network architecture which addresses the temporal collaborative ranking problem successfully capturing sequential data. Achieving state-of-the-art results, all of the five models provide unique and effective solutions to typical issues plaguing contemporary recommender systems.
\end{abstract}
\end{otherlanguage}

\newpage
\mytableofcontents
\frontmatter
\mainmatter
\include{bsc_english_contents}
% \cleardoublepage                          %fixes the position of bibliography in bookmarks
\phantomsection
\addcontentsline{toc}{chapter}{\bibname}  
\printbibliography

[1] Pan Li, Maofei Que, Zhichao Jiang, Yao Hu, and Alexander Tuzhilin. 2020. PURS: Personalized Unexpected Recommender System for Improving User Satisfaction. In Fourteenth ACM Conference on Recommender Systems (RecSys ’20), September 22–26, 2020, Virtual Event, Brazil. ACM, New York, NY, USA, 10 pages.
\begin{sloppypar}
https://doi.org/10.1145/3383313.3412238
\end{sloppypar}

[2] Jianing Sun, Wei Guo, Dengcheng Zhang, Yingxue Zhang, Florence Regol, Yaochen Hu, Huifeng Guo, Ruiming Tang, Han Yuan, Xiuqiang He, Mark Coates. 2020. A Framework for Recommending Accurate and Diverse Items Using Bayesian Graph Convolutional Neural Networks. In Proceedings of the 26rd ACM SIGKDD Conference on Knowledge Discovery and Data Mining (KDD’20), August 23-27, 2020, Virtual Event, CA, USA. ACM, New York, NY, USA, 10 pages. https://doi.org/10.1145/3397271.3401123

[3] Dalin Guo, Sofia Ira Ktena, Ferenc Huszar, Pranay Kumar Myana, Michael Kneier, Sourav Das, Wenzhe Shi, and Alykhan Tejani. 2020. Deep Bayesian Bandits: Exploring in Online Personalized Recommendations. In Fourteenth ACM Conference on Recommender Systems (RecSys ’20), September 22–26, 2020, Virtual Event, Brazil. ACM, New York, NY, USA, 6 pages. https://doi.org/10.1145/3383313.3412214

[4] Xian Wu, Suleyman Cetintas, Deguang Kong, Miao Lu, Jian Yang, and NiteshV. Chawla. 2020. Learning from Cross-Modal Behavior Dynamics withGraph-Regularized Neural Contextual Bandit. InProceedings of The WebConference 2020 (WWW ’20), April 20–24, 2020, Taipei, Taiwan.ACM, NewYork, NY, USA, 11 pages.
\newline
https://doi.org/10.1145/3366423.3380178

[5]Liwei Wu, Shuqing Li, Cho-Jui Hsieh, and James Sharpnack. 2020. SSE-PT:
Sequential Recommendation Via Personalized Transformer. In Fourteenth
ACM Conference on Recommender Systems (RecSys ’20), September 22–26,
2020, Virtual Event, Brazil. ACM, New York, NY, USA, 10 pages. https://doi.
org/10.1145/3383313.3412258

[6] (Li\&Tuzhilin, cited in Pan Li, Maofei Que, Zhichao Jiang, Yao Hu, and Alexander Tuzhilin. 2020.) 

[7] Personalized Glossary. https://www.dynamicyield.com/glossary/contextual-bandit/

[8] Rudin Cynthia. Basic Algorithms. Duke   
University.https://users.cs.duke.edu/~cynthia/CourseNotes/MABSlides.pdf

[9] Coenen Anna. 2016. How The New York Times is Experimenting with Recommendation Algorithms. NYTOpen. New York Times.
\begin{sloppypar}
https://open.nytimes.com/how-the-new-york-times-is-experimenting-with-recommendation-algorithms-562f78624d26
\end{sloppypar}

[10] Neural Graph Collaborative Filtering by Xiang et al. in 2019, and Graph Convolutional Neural Networks for Web-Scale Recommender Systems by Ying et al. in 2018.

\begin{sloppypar}
https://machinelearningmastery.com/a-gentle-introduction-to-the-bootstrap-method/
\end{sloppypar}


[11]Code and data are open-sourced at
https://github.com/wuliwei9278/SSE-PT

[12] Zhang, S, Yao, L., and Tay 2019. Deep Learning Based Recommender System: A Survey and New Perspectives” by, Y. published in ACM Survey 
\newline
https://dl-acm-org.libproxy.helsinki.fi/doi/abs/10.1145/3285029
\end{document}

